\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-08-27 ResponseToReview.tex

\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: ``For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and yield into domesticated rice (Kovach and McCouch 2008). Furthermore, in teosinte, many of the alleles underlying domestication phenotypes have been shown to be segregating in populations of teosinte, suggesting that natural variation in teosinte has played an important role in the domestication of maize (Gallavotti \emph{et al.} 2004; Sigmon and Vollbrecht 2010). While some effort has been invested into understanding how wild alleles behave in their domesticated relatives (\emph{e.g.} Bai and Lindhout 2007), very little is known about the role that alleles found most commonly in domesticates play in natural populations of their wild progenitors (Whitton \emph{et al.} 1997).}

8) Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte?\\

\res{We have moved the references cited here to more clearly reflect the papers' findings.  We have also added text describing the role of \emph{tb1} in the shade avoidance pathway and clarifying the manner in which this could affect fitness in teosinte}

\mbh{References need to follow text saying \emph{tb1} is important in the shade avoidance pathway and we need to add a sentence or two clarifying what it does in this pathway and the implications for teosinte fitness} \lev{the refs are in the sentences discussing the SAS in teo/grasses}

@@

9) Page 5, Lines 15 - 16: you say you ``aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic effects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions?\\

\res{Reviewer 2 is correct in noting that ssp. \emph{mexicana} is not as prevalent in our paper as ssp. \emph{parviglumis}. This fact is due to limitations in access to accessions of ssp. \emph{mexicana} as field collections for ssp. \emph{mexicana} are not nearly as extensive as those of ssp. \emph{parviglumis}. Additionally ssp. \emph{mexicana} is more difficult to cultivate in a greenhouse setting. We felt that it would be important to include ssp. \emph{mexicana} in our study, especially since it is known that gene flow occurs between ssp. \emph{mexicana} and landrace maize. We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.} \jri{just say we don't have as much seed from mexicana, which is true. it's just as easy to grow in the greenhouse. also I'd think of adding a sentence to the discussion mentioning mexicana. if it is in mexicana too, doesn't that suggest it's pretty old, given the divergence time between mexicana and parviglumis? or do we think it's in mexicana because of gene flow with parv or maize? this could be worth speculating on.}
\lev{Hmm. Well didn't Matt's paper on introgression support a significant lack of gene flow from maize to mex around domestication loci? So how likely would it be that it is in mex because of gene flow? So would it make more sense to argue that it's there because it is old and hasn't been lost? Would it make sense too because it was estimated to predate the domestication of maize? Maybe it is in parv at a higher frequency because it's 'doing' something, even though we didn't pick up an effect on tillering, and maybe it's just drifting in mex, which could explain why it is at a much lower frequency in mex?}

### 2014-08-27 manuscript.tex
edits
edits to response to reviewer 2; introduction & SAS sections

In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.  An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}. The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} and may therefore be crucial to the ecology of teosinte. In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} combined with a larger SNP dataset in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.

>>>>

In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.  An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}. The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} via changes in expression levels in response to shading, and may therefore be crucial to the ecology of teosinte. In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} combined with a larger SNP dataset in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.
